---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```

#read in clean data 
```{r}
long_clean_candy_data <- read_csv(here("clean_data/long_combined_clean_data.csv"))
```


#dimensions of clean data set
```{r}

dim(long_clean_candy_data)

```

# Question 1
# What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}

long_clean_candy_data %>% 
  select(ratings) %>% 
  summarise(total_num_ratings = sum(!is.na(ratings)))

```

The total number of candy ratings across the three years (excluding any NAs)
was 618139!

# Question 2
# What was the average age of people who are going out trick or treating?


```{r}

long_clean_candy_data %>% 
  select(are_you_going_actually_going_trick_or_treating_yourself) %>% 
  distinct()
```

```{r}

long_clean_candy_data %>% 
  filter(are_you_going_actually_going_trick_or_treating_yourself == "Yes") %>%
  summarise(avg_age_trick_treater = mean(age, na.rm = TRUE))

```

The average age for someone who went out trick or treating is 33.71 years (2dp).
