---
title: "R Notebook"
output: html_notebook
---

# .Rmd for analysing data

```{r}
head(decathlon_data_long)
```

# Who had the longest long jump seen in the data?

```{r}

decathlon_data_long %>% 
  select(athlete_name, event, ind_event_score, competition) %>% 
  filter(event == "long_jump") %>% 
  arrange(desc(ind_event_score)) %>% 
  head(1)

```

___Who had the longest long jump?___
The longest long jump seen in the data was that of Clay, with a distance of 
7.96m. This was also set in Olympic competition.

# What was the average 100m time in each competition?

___Average 100m time in decastar competition___

```{r}

decathlon_data_long %>% 
  select(competition, event, ind_event_score) %>% 
  filter(event == "x100m" &  competition == "Decastar") %>% 
  mutate(mean_100m_time = mean(ind_event_score))

```

The average 100m time set in the Decastar competition was 11.18 seconds, rounded
to 2 decimal places.

___Average 100m time in Olympic competition___

```{r}
decathlon_data_long %>% 
  select(competition, event, ind_event_score) %>% 
  filter(event == "x100m" &  competition == "OlympicG") %>% 
  mutate(mean_100m_time = mean(ind_event_score))
```

The average 100m time set in Olympic competition was 10.92 seconds, rounded to
2 decimal places. This was 0.26 seconds quicker than the mean time set in the
decastar competition.