---
title: "Dirty Data Project - Task 1"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  pdf_document: default
---
# Task Introduction
This data set included the results for two decathlon competitions; one from an
Olympic Games and the other from a Decastar competition. Included in the data
was information on the athletes names, their individual event scores and 
overall finishing position.

# Assumptions
My biggest assumption made was that all scores were entered in correctly.

# Cleaning Process
The first process to clean the data was to use `clean_names()` within the
janitor package to clean all column names. Then I removed the athlete names
as row names and moved them to their own column, while changing to lower case. 
This column was renamed from row_names to athlete_names.

Once the athlete names had their own column I used `pivot_longer()` on the
individual events so that I could analyse between different events later.

Finally, I renamed the columns of rank and score to final_position and
ind_event_score, respectively.

# Load Relevant Libraries

```{r}
library(readr)
library(tidyverse)
library(here)
```


# Read RDS file for analysing data

```{r}
decathlon_data_long <- read_rds("../clean_data/decathlon_data_long.RDS")
```

# Question 1

## Who had the longest long jump seen in the data?

```{r}

decathlon_data_long %>% 
  select(athlete_name, event, ind_event_score, competition) %>% 
  filter(event == "long_jump") %>% 
  arrange(desc(ind_event_score)) %>% 
  head(1)

```

___Who had the longest long jump?___
The longest long jump seen in the data was that of Clay, with a distance of 
7.96m. This was also set in Olympic competition.

# Question 2

## What was the average 100m time in each competition?

___Average 100m time in decastar competition___

```{r}

decathlon_data_long %>% 
  select(competition, event, ind_event_score) %>% 
  filter(event == "x100m" &  competition == "Decastar") %>% 
  mutate(mean_100m_time = mean(ind_event_score))

```

The average 100m time set in the Decastar competition was 11.18 seconds, rounded
to 2 decimal places.

___Average 100m time in Olympic competition___

```{r}

decathlon_data_long %>% 
  select(competition, event, ind_event_score) %>% 
  filter(event == "x100m" &  competition == "OlympicG") %>% 
  summarise(mean_100m_time = mean(ind_event_score))

```

The average 100m time set in Olympic competition was 10.92 seconds, rounded to
2 decimal places. This was 0.26 seconds quicker than the mean time set in the
decastar competition.

# Question 3

## Who had the highest total points across both competitions?

```{r}

decathlon_data_wide <- decathlon_data_long %>%
  pivot_wider(names_from = event, 
              values_from = ind_event_score)

decathlon_data_wide %>% 
  group_by(athlete_name) %>%                         
  summarise_at(vars(points),            
               list(points_total = sum)) %>% 
  arrange(desc(points_total)) %>% 
  head(1)

```

The decathlete with the greatest points total across both competitions was 
Roman Sebrle, with a combined points total of 17110 points! He was only 168
points ahead of Will Clay who came second across both competitions. For example,
that equates to roughly 0.65 seconds of a difference in 100m times, fine margins!

# Question 4

## What was the shot-put scores for the top three competitors in each competition?

___Top 3 shot-put scores in Decastar competition___

```{r}

decathlon_data_long %>% 
  select(athlete_name, competition, event, ind_event_score) %>% 
  filter(competition == "Decastar" & event == "shot_put") %>% 
  arrange(desc(ind_event_score)) %>% 
  head(3)

```

The top three shot-put scores in the Decastar competition were from Yurkov, 
Sebrle and Karpov with scores of 15.19m, 14.83m and 14.77m respectively.

___Top 3 shot-put scores in Olympic competition___

```{r}

decathlon_data_long %>% 
  select(athlete_name, competition, event, ind_event_score) %>% 
  filter(competition == "OlympicG" & event == "shot_put") %>% 
  arrange(desc(ind_event_score)) %>% 
  head(3)

```

The top 3 shot-put scores in Olympic competition came from Sebrle, Karpov and
Macey with scores of 16.36m, 15.93m and 15.73m respectively. 

# Question 5

## What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

___Average overall score with a 400m time < 50 seconds___

```{r}

decathlon_data_long %>% 
  filter(event == "x400m" & ind_event_score < 50.00) %>% 
  summarise(avg_score = mean(points))

```

The average score for athletes across both competitions who ran the 400m in 
less than 50 seconds was 8120.483.

___Average overall score with a 400m time > 50 seconds___

```{r}

decathlon_data_long %>% 
  filter(event == "x400m" & ind_event_score > 50.00) %>% 
  summarise(avg_score = mean(points))

```

The average score across both competitions when an athlete ran the 400m in 
greater than 50 seconds was 7727.167. This gives a difference in 393.316 points
in an athletes average total score depending on whether their 400m was greater
or less than 50 seconds. 