---
title: "R Notebook"
output: html_notebook
---

# .Rmd for analysing data

```{r}
head(decathlon_data_long)
```

# Who had the longest long jump seen in the data?

```{r}

decathlon_data_long %>% 
  select(athlete_name, event, ind_event_score, competition) %>% 
  filter(event == "long_jump") %>% 
  arrange(desc(ind_event_score)) %>% 
  head(1)

```

___Who had the longest long jump?___
The longest long jump seen in the data was that of Clay, with a distance of 
7.96m. This was also set in Olympic competition.

# What was the average 100m time in each competition?

___Average 100m time in decastar competition___

```{r}

decathlon_data_long %>% 
  select(competition, event, ind_event_score) %>% 
  filter(event == "x100m" &  competition == "Decastar") %>% 
  mutate(mean_100m_time = mean(ind_event_score))

```

The average 100m time set in the Decastar competition was 11.18 seconds, rounded
to 2 decimal places.

___Average 100m time in Olympic competition___

```{r}
decathlon_data_long %>% 
  select(competition, event, ind_event_score) %>% 
  filter(event == "x100m" &  competition == "OlympicG") %>% 
  mutate(mean_100m_time = mean(ind_event_score))
```

The average 100m time set in Olympic competition was 10.92 seconds, rounded to
2 decimal places. This was 0.26 seconds quicker than the mean time set in the
decastar competition.

# Who had the highest total points across both competitions?

```{r}

decathlon_data_wide <- decathlon_data_long %>%
  pivot_wider(names_from = event, 
              values_from = ind_event_score)

decathlon_data_wide %>% 
  group_by(athlete_name) %>%                         
  summarise_at(vars(points),            
               list(points_total = sum)) %>% 
  arrange(desc(points_total)) %>% 
  head(1)

```

The decathlete with the greatest points total across both competitions was 
Roman Sebrle, with a combined points total of 17110 points! He was only 168
points ahead of Will Clay who came second across both competitions. For example,
that equates to roughly 0.65 seconds of a difference in 100m times, fine margins!

# What was the shot-put scores for the top three competitors in each competition?

___Top 3 shot-put scores in Decastar competition___

```{r}

decathlon_data_long %>% 
  select(athlete_name, competition, event, ind_event_score) %>% 
  filter(competition == "Decastar" & event == "shot_put") %>% 
  arrange(desc(ind_event_score)) %>% 
  head(3)

```

The top three shot-put scores in the Decastar competition were from Yurkov, 
Sebrle and Karpov with scores of 15.19m, 14.83m and 14.77m respectively.

___Top 3 shot-put scores in Olympic competition___

```{r}
decathlon_data_long %>% 
  select(athlete_name, competition, event, ind_event_score) %>% 
  filter(competition == "OlympicG" & event == "shot_put") %>% 
  arrange(desc(ind_event_score)) %>% 
  head(3)

```

The top 3 shot-put scores in Olympic competition came from Sebrle, Karpov and
Macey with scores of 16.36m, 15.93m and 15.73m respectively. 